### 2018-10-29 Chapters/TDHeff.tex

First we will start we a general Hamiltonian of the type:

\begin{equation}
\hat{H} = -\hat{T} + \text{U}\hat{D}
\end{equation}

Where $\hat{D} = \sum_{i=1}^M \hat{n}_{i\uparrow}\hat{n}_{i\downarrow}$ is the doublon number operator and $\hat{T}$ is any hopping operator. The only assumption we will make about $\hat{T}$ is that it is proportional to any power of $\hat{p}$ in which case we can apply the Peierls substitution (\ref{AP3A}) to introduce to take into account the effect of the electromagnetic perturbation in the lattice. The strength of the on-site interaction $\text{U}$ is much larger than the hopping amplitude, therefore, in a half filling system the zero double occupancies subspace $d=0$ can be taken as the low energy subspace in which the effective Hamiltonian will act.

In terms of on-site creation and annihilation operators we can write the hopping operator as $\hat{T} = \sum_{i,j, \sigma, \sigma'} t_{ij}^{\sigma \sigma'} \hat{c}_{i \sigma}^\dagger \hat{c}_{j \sigma'}$. In the simplest case, for example, the sum would be restricted to nearest neighbors and the hopping amplitude would be diagonal in the spin state. With this notation, in presence of a vector potential $\vec{A}(t)$ (which we assume to not vary noticeably in the scale of the lattice) the Peierls substitution leads to an extra time dependent phase in the hopping amplitude:

\begin{equation}
t_{ij}^{\sigma \sigma'}(t) = t_{ij}^{\sigma \sigma'} e^{ie\bs{R}_{ij}\bs{A}(t)}
\end{equation}

Where $\bs{A}$ is the vector potential and $\bs{R}_{ij} = \bs{R}_i-\bs{R}_j$. We write the electric field as $\bs{E}(t) = \frac{1}{2}(\vec{E}e^{-i\omega t}+\vec{E}^*e^{i\omega t})$, where $\vec{E} = E_0\hat{e}$ and $\hat{e} = \frac{1}{\sqrt{1+\lambda_{POL}^2}}(\hat{e}_x+i\lambda_{POL}\hat{e}_y)$ is the polarization vector and $\lambda_{POL} = 0, \pm 1$ for plane polarized, right handed and left handed circular polarized field respectively. The vector potential takes the form $\bs{A}(t) = \frac{1}{2}(\vec{A}e^{-i\omega t} + \vec{A}^* e^{i\omega t})$, with $\vec{A} = \frac{iE_0}{\omega}\hat{e}$.

Let us define $e\bs{R}_{ij}\vec{A} = \alpha_{ij} e^{i \theta_{ij}}$, with $\alpha_{ij} = \pm|e\bs{R}_{ij}\vec{A}|$ in such a way that:

\begin{align}
\alpha_{ij} &= -\alpha_{ji} \label{alphaSym} \\
\theta_{ij} &= \theta_{ji} \label{thetaSym}
\end{align}

and $\theta_{ij} \in \left[0,\pi\right)$. Then we can apply the Jacobi–Anger expansion:

\begin{align*}
t_{ij}^{\sigma \sigma'}(t) = t_{ij}^{\sigma \sigma'}e^{ie\bs{R}_{ij}\bs{A}(t)} = t_{ij}^{\sigma \sigma'}e^{i\alpha_{ij} \cos(\omega t - \theta_{ij})} = t_{ij}^{\sigma \sigma'}\sum_m e^{i(\frac{\pi}{2}-\theta_{ij})m} \mathcal{J}_m(\alpha_{ij}) e^{im\omega t} = \sum t_{ij,m}^{\sigma \sigma'} e^{im\omega t}
\end{align*}

Where we defined

\begin{equation}
\label{HoppAmpFourier}
t_{ij,m}^{\sigma \sigma'} = t_{ij}^{\sigma \sigma'} e^{i(\frac{\pi}{2}-\theta_{ij})m} \mathcal{J}_m(\alpha_{ij})
\end{equation}

Which is the \textit{m}th Fourier mode of the hopping term and $\mathcal{J}_m(x)$ is the \textit{m}th Bessel function \cite{Kitamura2017}. Correspondingly we can write $\hat{T}(t) = \sum_m \hat{T}_m e^{im \omega t}$ where $\hat{T}_m$ is the sum of all the \textit{m}th Fourier mode of the hopping terms. We can further decompose the hopping operator into:

\begin{equation}
\hat{T}(t) = \sum_m (\hat{T}_{-1,m}+\hat{T}_{0,m}+\hat{T}_{1,m})e^{im\omega t}
\end{equation}

Where $\hat{T}_{dm}(t)$ changes the doublon number by $d$, for example, if $\hat{P}_d$ is the projection operator into the subspace with doublon number $d$, then $\hat{T}_{dm}(t) = \sum_i \hat{P}_{i+d}\hat{T}_{m}(t)\hat{P}_i$.

In order to derive the form of the effective Hamiltonian let us introduce a time dependent unitary transformation $\hat{U}(t) = e^{-i\hat{S}(t)}$. The transformed Hamiltonian is:

\begin{equation}
\hat{H}'(t) = e^{i\hat{S}(t)} \hat{H}(t) e^{-i\hat{S}(t)} - e^{i\hat{S}(t)} id_t e^{-i\hat{S}(t)}
\end{equation}

We perform the unitary transformation perturbatively in the hopping operator, we can formally write $\hat{T}(t) = \eta \hat{T}(t)$, where $\eta$ will play the role of a bookkeeping parameter in the perturbative expansion. We expand $\hat{S}(t) = \sum_\nu \eta^\nu \hat{S}^{(\nu)}(t)$ and $\hat{H}'(t) = \sum_\nu \eta^\nu \hat{H}'^{(\nu)}(t)$. In order for the new Hamiltonian to be periodic we impose the unitary transformation to have the periodicity of the Hamiltonian, so that we can expand $\hat{S}^{(\nu)}(t) = \sum_m e^{im\omega t}\hat{S}^{(\nu)}_m$. Additionally we can impose the transformed Hamiltonian to be block diagonal in the doublon number $d$. With these conditions the unitary transformation can be uniquely determined if we impose that $\hat{S}(t)$ does not contain block-diagonal terms, i.e. we can write:

\begin{equation}
\hat{S}^{(\nu)}(t) = \sum_{d \neq 0} \sum_m \eta^\nu \hat{S}^{(\nu)}_{d,m} e^{im\omega t}
\end{equation}

where $\hat{S}^{(\nu)}_{d,m}$ changes the double occupancy number by $d$.

Now we can use the identity:

\begin{equation}
id_t e^{-i\hat{S}(t)} = \sum_n \frac{1}{(n+1)!}\text{ad}_{-i\hat{S}(t)}^n (d_t \hat{S}(t))e^{-i\hat{S}(t)}
\end{equation}

Derived in appendix \ref{AP3B}, where $\text{ad}_A(B) = [A,B]$ to rewrite the transformed Hamiltonian as:

\begin{equation}
\hat{H}'(t) = e^{i\hat{S}(t)} \left( \hat{H}(t) - \sum_n \frac{1}{(n+1)!}\text{ad}_{-i\hat{S}(t)}^n (d_t \hat{S}(t)) \right) e^{-i\hat{S}(t)}
\end{equation}

Using the Baker–Campbell–Hausdorff expansion we can write the transformed Hamiltonian as:

\begin{equation}
\label{PertFull}
\hat{H}'(t) = \sum_m \frac{1}{m!} \text{ad}_{i\hat{S}(t)}^m \left( \hat{H}(t) - \sum_n \frac{1}{(n+1)!}\text{ad}_{-i\hat{S}(t)}^n (d_t \hat{S}(t)) \right)
\end{equation}

Now, in this expression we have to expand $\hat{S}(t) = \sum_\nu \eta^\nu \hat{S}^{(\nu)}(t)$ and $\hat{H}'(t) = \sum_\nu \eta^\nu \hat{H}'^{(\nu)}(t)$ and determine $\hat{S}^{(\nu)}(t)$ iteratively in $\nu$ so that $\hat{H}'^{(\nu)}(t)$ is diagonal in the doublon number. Notice that we do not expand $\hat{H}(t)$ as an infinite series since $\hat{H}(t) = -\eta \hat{T}(t) + \text{U}\hat{D}$. Now, notice also that if we take $i\hat{S}^{(0)}(t)=0$ in \ref{PertFull}, then, to zeroth order in $\eta$ only terms in $m=0$ and $n=0$ contribute obtaining:

\begin{equation}
\hat{H}'^{(0)}(t) = \text{U}\hat{D}
\end{equation}

Which is indeed diagonal in the doublon number, and therefore we can take $i\hat{S}^{(0)}(t)=0$. With this equating terms of the same order in \ref{PertFull} is greatly simplified, since terms with high $m,n$ will correspond to terms with high $\eta$. Since we are only interested in orders up to second order, we can truncate \ref{PertFull} up to $m=2$:

\begin{align}
\label{PertTrunc}
\hat{H}'(t) &=  \hat{H}(t) - \sum_n \frac{1}{(n+1)!}\text{ad}_{-i\hat{S}(t)}^n (d_t \hat{S}(t)) + [i\hat{S}(t), \hat{H}(t) - \sum_n \frac{1}{(n+1)!}\text{ad}_{-i\hat{S}(t)}^n (d_t \hat{S}(t))] + \nonumber \\
&+ \frac{1}{2} [i\hat{S}(t),[i\hat{S}(t), \hat{H}(t) - \sum_n \frac{1}{(n+1)!}\text{ad}_{-i\hat{S}(t)}^n (d_t \hat{S}(t)) ]]
\end{align}

This equation holds up to order $\eta^2$. In first order we obtain:

\begin{equation}
\label{1stO}
\hat{H}'^{(1)}(t) = -\hat{T}(t) - d_t\hat{S}^{(1)}(t) + \left[ i\hat{S}^{(1)}(t), \text{U}\hat{D} \right]
\end{equation}

Now we expanding in $m,d$ and use $\left[ \hat{D}, \hat{S}^{(\nu)}_{dm} \right] = d\hat{S}^{(\nu)}_{dm}$ because $\hat{S}^{(\nu)}_{dm}$ changes the doublon number by $d$:

\begin{equation}
\hat{H}'^{(1)}(t)=-\hat{T}(t)-\sum_{d\neq 0}\sum_m (\text{U}d+m\omega) i\hat{S}^{(1)}_{dm} e^{im\omega t}
\end{equation}

Therefore:

\begin{align}
i\hat{S}^{(1)}_d(t) &= -\sum_m \frac{\hat{T}_{d,m}}{\text{U}d+m\omega}e^{im\omega t} \label{1stOSpin}\\
\hat{H}'^{(1)}(t) &= -\sum_m \hat{T}_{0,m}(t)e^{im\omega t} \label{1stOH}
\end{align}

To second order we find:

\begin{align}
\hat{H}'^{(2)}(t) &= - d_t\hat{S}^{(2)}(t) - \frac{1}{2}\left[-i\hat{S}^{(1)}(t), d_t\hat{S}^{(1)}(t) \right] + \left[i\hat{S}^{(1)}(t), -\hat{T}(t)-\frac{1}{2}d_t\hat{S}^{(1)}(t) \right] +\nonumber \\
&+ \left[i\hat{S}^{(2)}(t), \text{U}\hat{D} \right] + \frac{1}{2} \left[i\hat{S}^{(1)}(t), \left[i\hat{S}^{(1)}(t), \text{U}\hat{D} \right] \right] = \nonumber \\
&= \left[i\hat{S}^{(2)}(t), \text{U} \hat{D} \right] - \left[ i\hat{S}^{(1)}(t), \hat{T}(t) \right] + \frac{1}{2}\left[ i\hat{S}^{(1)}(t), \left[ i\hat{S}^{(1)}(t), \text{U}\hat{D} \right] \right] - d_t\hat{S}^{(2)}(t)
\end{align}

Using \ref{1stO} and using $\left[ i\hat{S}^{(1)}(t), d_t\hat{S}^{(1)}(t)\right] = 0$ due to different Fourier modes of the hopping operator being equal up to a constant (explain better), we can rewrite:

\begin{equation}
\hat{H}'^{(2)}(t) = \left[i\hat{S}^{(2)}(t), \text{U} \hat{D} \right] - \frac{1}{2}\left[ i\hat{S}^{(1)}(t), \hat{T}(t) - \hat{H}'^{(1)}(t)\right] - d_t\hat{S}^{(2)}(t)
\end{equation}

Using \ref{1stOSpin} and \ref{1stOH}, the middle term is:

\begin{align*}
&\left[ i\hat{S}^{(1)}(t), \hat{T}(t) - \hat{H}'^{(1)}(t)\right] = -\left[\sum_m \left( \frac{\hat{T}_{1m}}{\text{U}+m\omega} - \frac{\hat{T}_{-1m}}{\text{U}-m\omega} \right)e^{im \omega t}, \sum_n \left( \hat{T}_{-1n} + 2\hat{T}_{0n} + \hat{T}_{1n} \right) e^{in\omega t} \right] \\
&= -\sum_{mn} \left\{ \frac{2\left[\hat{T}_{1m}, \hat{T}_{0n} \right]}{\text{U}+m\omega} + \frac{\left[\hat{T}_{1m}, \hat{T}_{-1n} \right]}{\text{U}+m\omega} - \frac{\left[\hat{T}_{-1m}, \hat{T}_{1n} \right]}{\text{U}-m\omega} - \frac{2\left[\hat{T}_{-1m}, \hat{T}_{0n} \right]}{\text{U}-m\omega} \right\} e^{i(m+n)\omega t} \\
&= -\sum_{mn} \left\{ \frac{2\left[\hat{T}_{1n}, \hat{T}_{0(m-n)} \right]}{\text{U}+n\omega} + \frac{\left[\hat{T}_{1n}, \hat{T}_{-1(m-n)} \right]}{\text{U}+n\omega} - \frac{\left[\hat{T}_{-1n}, \hat{T}_{1(m-n)} \right]}{\text{U}-n\omega} - \frac{2\left[\hat{T}_{-1n}, \hat{T}_{0(m-n)} \right]}{\text{U}-n\omega} \right\} e^{im\omega t}
\end{align*}

Altogether:

\begin{align*}
\hat{H}'^{(2)}(t) &= \sum_{mn} \left\{ \frac{\left[\hat{T}_{1n}, \hat{T}_{0(m-n)} \right]}{\text{U}+n\omega} + \frac{\left[\hat{T}_{1n}, \hat{T}_{-1(m-n)} \right]}{2(\text{U}+n\omega)} - \frac{\left[\hat{T}_{-1n}, \hat{T}_{1(m-n)} \right]}{2(\text{U}-n\omega)} - \frac{\left[\hat{T}_{-1n}, \hat{T}_{0(m-n)} \right]}{\text{U}-n\omega} \right\} e^{im\omega t} \\
&-\sum_{d\neq 0}\sum_m (\text{U}d+m\omega) i\hat{S}^{(2)}_{dm} e^{im\omega t}
\end{align*}

We obtain:

\begin{align}
i\hat{S}^{(2)}_{dm} &= \sum_n \frac{\left[ \hat{T}_{dn}, \hat{T}_{0(m-n)} \right]}{(\text{U}d+n\omega)(\text{U}d+m\omega)} \label{2ndOSpin}\\
\hat{H}'^{(2)}(t) &= \frac{1}{2}\sum_{mn} \left( \frac{\left[\hat{T}_{1n}, \hat{T}_{-1(m-n)} \right]}{\text{U}+n\omega} - \frac{\left[\hat{T}_{-1n}, \hat{T}_{1(m-n)} \right]}{\text{U}-n\omega} \right) e^{im\omega t} \label{2ndOH}
\end{align}

Now the effective Hamiltonian acts on the subspace of $\hat{P}_0$, therefore we are only interested on $\hat{P}_0 \hat{H}'^{(2)}(t) \hat{P}_0$ (notice that $\hat{P}_0 \hat{H}'^{(1)}(t) \hat{P}_0 = 0$). In this subspace we can write:

\begin{align}
\hat{H}_{\text{eff}}(t) &= \hat{P}_0\hat{H}'^{(2)}(t)\hat{P}_0 = -\frac{1}{2}\sum_{mn} \left( \frac{\hat{P}_0  \hat{T}_{-1(m-n)}\hat{T}_{1n}\hat{P}_0}{\text{U}+n\omega} + \frac{\hat{P}_0 \hat{T}_{-1n} \hat{T}_{1(m-n)} \hat{P}_0}{\text{U}-n\omega} \right) e^{im\omega t} \nonumber \\
&= -\frac{1}{2}\sum_{mn} \left\{ \frac{\hat{P}_0  (\hat{T}_{-1(m-n)}\hat{T}_{1n} + \hat{T}_{-1-n}\hat{T}_{1(m+n)})\hat{P}_0}{\text{U}+n\omega} \right\} e^{im\omega t}
\end{align}

Now, since the hopping operators act on the subspace $d=0$ and must remain within that subspace, therefore $\hat{P}_0 \hat{T}_{-1a} \hat{T}_{1b} \hat{P}_0$ will be a sum of all possible hoppings between two different sites:

\begin{align*}
\hat{P}_0 \hat{T}_{-1a} \hat{T}_{1b} \hat{P}_0 = \sum_{i,j, \sigma_1, \sigma_2, \sigma_3, \sigma_4} t_{ij,a}^{\sigma_1 \sigma_2} t_{ji,b}^{\sigma_3 \sigma_4} \hat{c}_{i \sigma_1}^\dagger \hat{c}_{j \sigma_2} \hat{c}_{j \sigma_3}^\dagger \hat{c}_{i \sigma_4}
\end{align*}

Where $a$ and $b$ are any two Fourier modes. Inserting this into \ref{HeffSimplified} and using the definition of $t_{ij,m}^{\sigma \sigma'} = t_{ij}^{\sigma \sigma'} e^{i(\frac{\pi}{2}-\theta_{ij})m} \mathcal{J}_m(\alpha_{ij})$ and using that $\theta_{ji} = \theta_{ij}$ and $\alpha_{ji} = -\alpha_{ij}$ and the properties $\mathcal{J}_{-m}(x) = (-1)^m\mathcal{J}_m(x)$ and $\mathcal{J}_m(-x) = (-1)^m\mathcal{J}_m(x)$ we can write \ref{HeffSimplified} as:

\begin{align}
&\hat{H}_{\text{eff}}(t) = \hat{P}_0\hat{H}'^{(2)}(t)\hat{P}_0 = - \frac{1}{2}\sum_{mn} \sum_{i,j, \sigma_1, \sigma_2, \sigma_3, \sigma_4}\hat{c}_{i \sigma_1}^\dagger \hat{c}_{j \sigma_2} \hat{c}_{j \sigma_3}^\dagger \hat{c}_{i \sigma_4} \frac{t_{ij,m-n}^{\sigma_1 \sigma_2} t_{ji,n}^{\sigma_3 \sigma_4} + t_{ij,-n}^{\sigma_1 \sigma_2} t_{ji,m+n}^{\sigma_3 \sigma_4}}{\text{U}+n\omega} e^{im\omega t} \nonumber \\
&= - \frac{1}{2}\sum_{mn} \sum_{i,j, \sigma_1, \sigma_2, \sigma_3, \sigma_4}\hat{c}_{i \sigma_1}^\dagger \hat{c}_{j \sigma_2} \hat{c}_{j \sigma_3}^\dagger \hat{c}_{i \sigma_4} t_{ij}^{\sigma_1 \sigma_2} t_{ji}^{\sigma_3 \sigma_4} e^{i(\frac{\pi}{2}-\theta_{ij})m} \frac{ \mathcal{J}_{m-n}(\alpha_{ij}) \mathcal{J}_{n}(\alpha_{ji}) + \mathcal{J}_{-n}(\alpha_{ij}) \mathcal{J}_{m+n}(\alpha_{ji})}{\text{U}+n\omega} e^{im\omega t} \nonumber \\
&= - \frac{1}{2}\sum_{mn} \sum_{i,j, \sigma_1, \sigma_2, \sigma_3, \sigma_4}\hat{c}_{i \sigma_1}^\dagger \hat{c}_{j \sigma_2} \hat{c}_{j \sigma_3}^\dagger \hat{c}_{i \sigma_4} t_{ij}^{\sigma_1 \sigma_2} t_{ji}^{\sigma_3 \sigma_4} e^{i(\frac{\pi}{2}-\theta_{ij})m} (-1)^m \frac{ \mathcal{J}_{n-m}(\alpha_{ij}) \mathcal{J}_{n}(\alpha_{ij}) + \mathcal{J}_{n}(\alpha_{ij}) \mathcal{J}_{m+n}(\alpha_{ij})}{\text{U}+n\omega} e^{im\omega t} \nonumber \\
&= - \frac{1}{2} \sum_{i,j, \sigma_1, \sigma_2, \sigma_3, \sigma_4}\hat{c}_{i \sigma_1}^\dagger \hat{c}_{j \sigma_2} \hat{c}_{j \sigma_3}^\dagger \hat{c}_{i \sigma_4} t_{ij}^{\sigma_1 \sigma_2} t_{ji}^{\sigma_3 \sigma_4} \mathcal{M}(\alpha_{ij}, \text{U}, \omega, t) \label{HeffSimplified}
\end{align}

Where we defined:

\begin{equation}
\mathcal{M}(\alpha_{ij}, \text{U}, \omega, t) = \sum_{mn}e^{-i(\frac{\pi}{2}+\theta_{ij})m} \left\{ 
    \frac{\mathcal{J}_{n-m}(\alpha_{ij})\mathcal{J}_{n}(\alpha_{ij}) + \mathcal{J}_{n}(\alpha_{ij})\mathcal{J}_{m+n}(\alpha_{ij})}{\text{U}+n\omega} \right\}e^{im\omega t}
\end{equation}

In the following sections we will apply what we have derived here for different models of the hopping operator.