### 2018-12-03 Chapters/Chapter3.tex
moar

\chapter{Perturbing the lattice}

In this chapter we will derive an effective spin Hamiltonian for different models in the presence of a high frequency electromagnetic perturbation. The goal is to manipulate $J$ and $\bs{D}_{ij}$ in \ref{HeffNOEMF} separately. This has already been done for the case of Mott insulators (\cite{Mentink2015}, \cite{Kitamura2017}) where the starting Hamiltonian is the Hubbard model:

\begin{equation*}
\hat{H} = -t_0\sum_{\langle i,j \rangle, \sigma} \hat{c}_{i \sigma}^\dagger \hat{c}_{j \sigma} + \text{U} \sum_{i=1}^M \hat{n}_{i\uparrow}\hat{n}_{i\downarrow}
\end{equation*}

In this model, since SOI is not taken into account, only the exchange interaction $J$ plays a role. The electromagnetic perturbation is then introduced via the Peierls substitution. We reproduce the result in section \ref{Section3Hubbard}. In the following sections we extend the analysis to different systems with spin orbit interaction thereby modulating $\bs{D}_{ij}$ or other spin interactions. The simplest model with spin orbit interaction is \ref{Hubbard}:

\begin{equation*}
\hat{H} = -\sum_{\langle i,j \rangle, \sigma, \sigma'}(\delta_{\sigma, \sigma'} t_{0} + \bs{\Delta}_{ij} \bs{\sigma}_{\sigma, \sigma'})\hat{c}_{i \sigma}^\dagger \hat{c}_{j \sigma'} + \text{U} \sum_{i=1}^M \hat{n}_{i\uparrow}\hat{n}_{i\downarrow}
\end{equation*}

Which is the same Hubbard model plus a spin dependent hopping term, which represents Rashba SOI for $\bs{\Delta}_{ij} = i\Delta_R(R_{ij}^y, -R_{ij}^x, 0)$ or Dresselhaus SOI for $\bs{\Delta}_{ij} = i\Delta_R(R_{ij}^x, -R_{ij}^y, 0)$. In section \ref{Section3HubbardSOI} we show how the exchange interaction and the DMI couplings are modulated in presence of an electromagnetic field.

In the case of graphene or other coplanar honeycomb lattice structures, nearest neighbor Rashba terms are not allowed by symmetry, and only next nearest neighbor (NNN) spin conserving terms contribute \cite{Kane2005}, this term opens a gap in the band structure describing a topological insulator. For this case we examine the Kane-Mele-Hubbard model. The effective spin model of this Hamiltonian contains the usual exchange interaction and DMI and, in addition, it exhibits anisotropic exchange interaction arising from the intrinsic SOI term. This new term has already been obtained in \cite{Rachel2010} without taking into account any electromagnetic field.

Non-coplanar honeycomb lattice may exhibit intrinsic Rashba interactions between NNN \cite{Liu2011}, this, together with a NNN hopping (which would arise from direct overlapping of the next nearest neighbor Wannier functions or from second order processes) will lead to NNN DMI.

\begin{section}{Time dependent effective Hamiltonian}
\label{SectionTDHeff}
\input{Chapters/TDHeff}

\end{section}

\begin{section}{Hubbard model without SOI}
\label{Section3Hubbard}
We start by considering the simplest hopping operator, as in the Hubbard model:

\begin{equation}
\hat{H} = -t_0\sum_{\langle i,j \rangle, \sigma} \hat{c}_{i \sigma}^\dagger \hat{c}_{j \sigma} + \text{U} \sum_{i=1}^M \hat{n}_{i\uparrow}\hat{n}_{i\downarrow}
\end{equation}

Which is the Hamiltonian \ref{Hubbard} without DMI interaction. We see that the hopping amplitudes introduced in the previous section take the form $t_{ij}^{\sigma \sigma'} = \delta_{\sigma \sigma'} t_0$ for $i,j$ being nearest neighbors, and $t_{ij}^{\sigma \sigma'} = 0$ otherwise. With this we can directly apply \ref{HeffSimplified} to obtain:

\begin{equation}
\hat{H}_{\text{eff}}(t) = -\frac{t_0^2}{2} \sum_{\langle i,j \rangle, \sigma, \sigma'} \hat{c}_{i \sigma}^\dagger \hat{c}_{j \sigma} \hat{c}_{j \sigma'}^\dagger \hat{c}_{i \sigma'} \mathcal{M}(\alpha_{ij}, \text{U}, \omega, t)
\end{equation}

Now, introducing the spin operators \ref{SpinOperatorInv1} and \ref{SpinOperatorInv2} and summing over the spin states as in \ref{SpinRel1}:

\begin{align*}
\sum_{\sigma, \sigma'} \hat{c}_{i \sigma}^\dagger \hat{c}_{j \sigma} \hat{c}_{j \sigma'}^\dagger \hat{c}_{i \sigma'} = \sum_{\sigma, \sigma'} \left( \frac{\delta_{\sigma \sigma'}}{2} + \bs{S}_i\bs{\sigma}_{\sigma' \sigma} \right) \left( \frac{\delta_{\sigma \sigma'}}{2} - \bs{S}_j\bs{\sigma}_{\sigma \sigma'} \right) = -2\bs{S}_i \bs{S}_j
\end{align*}

Where we neglected the constant term. The effective spin Hamiltonian is thus:

\begin{equation}
\hat{H}_{\text{eff}}(t) = \sum_{\langle i,j \rangle} J_{ij}(t) \bs{S}_i \bs{S}_j
\end{equation}

With

\begin{align}
J_{ij}(t) &= t_0^2 \mathcal{M}(\alpha_{ij}, \text{U}, \omega, t) \nonumber \\
&=\sum_{mn} t_0^2 e^{-i(\frac{\pi}{2}+\theta_{ij})m}\left(\frac{\mathcal{J}_{n-m}(\alpha_{ij})\mathcal{J}_{n}(\alpha_{ij})+\mathcal{J}_{n}(\alpha_{ij})\mathcal{J}_{n+m}(\alpha_{ij})}{\text{U}+n\omega} \right) e^{im\omega t} \label{Jij1}
\end{align}

After time average this reduces to the $m=0$ term:

\begin{equation}
\mathcal{M}(\alpha_{ij}, \text{U}, \omega, t) \approx \sum_{n} 2 \frac{\mathcal{J}_n(\alpha_{ij})^2}{\text{U}+n\omega}
\end{equation}

For $\omega>>U$ we can truncate this to the three smallest values of $n$, i.e. $n=0, \pm 1$. We can also use that $\alpha_{ij} << 1$ because $\alpha_{ij}$ is proportional to $\vec{A}$ which is proportional to $\omega^{-1}$. Therefore we can use $\mathcal{J}_n(x) \approx x^n \text{ for } n>0 \text{ and } x << 1$ to obtain:

\begin{equation}
\label{MFactorApprox}
\mathcal{M}(\alpha_{ij}, \text{U}, \omega, t) \approx 2 \left(\frac{\alpha_{ij}^2}{\text{U}+\omega} +\frac{1}{\text{U}} +\frac{\alpha_{ij}^2}{\text{U}-\omega} \right)
\end{equation}

And the exchange interaction coupling becomes:

\begin{equation}
\label{Jij2}
J_{ij} \approx J_{ij}^0 + 2t_0^2 \alpha_{ij}^2 \left( \frac{1}{\text{U}+\omega} + \frac{1}{\text{U}-\omega} \right)
\end{equation}

The lattice structure is contained in $\alpha_{ij} = \pm|e\bs{R}_{ij}\vec{A}|$. For a square lattices aligned with the coordinate system so that $\vec{a}_1=\hat{e}_x$ and $\vec{a}_2=\hat{e}_y$, we will have $\bs{R}_{ij} = \pm a\hat{e}_x,\pm a\hat{e}_y$, where $a$ is the lattice constant. Using $\vec{A}=\frac{iE_0}{\omega\sqrt{1+\lambda^2}}(\hat{e}_x+i\lambda\hat{e}_y)$ we have:

\begin{equation}
\alpha_{ij} = \begin{cases}
             \pm \frac{eaE_0}{\omega \sqrt{1+\lambda^2}} = \pm \frac{\mathcal{E}}{\sqrt{1+\lambda^2}},  & \text{for } \bs{R}_{ij} = \pm \hat{e}_x \\
             \pm \lambda\frac{eaE_0}{\omega \sqrt{1+\lambda^2}} = \pm \frac{\lambda \mathcal{E}}{\sqrt{1+\lambda^2}},  & \text{for } \bs{R}_{ij} = \pm \hat{e}_y
       \end{cases} \quad
\end{equation}

Where $\mathcal{E} = \frac{eaE_0}{\omega}$.

For plane polarized light ($\lambda = 0$) the exchange interaction only changes in the direction of the polarization, that is:

\begin{equation}
J_{ij}^{PP} = \begin{cases}
		J_{ij}^0 + 2t_0^2 \mathcal{E}^2 \left( \frac{1}{\text{U}+\omega} + \frac{1}{\text{U}-\omega} \right) & \text{for } \bs{R}_{ij} = \pm \hat{e}_x \\
J_{ij}^0 & \text{for } \bs{R}_{ij} = \pm \hat{e}_y
\end{cases} \quad 
\end{equation}

For circular polarized light ($\lambda=\pm1$) in this approximation the exchange interaction changes in the same way in all the directions:

\begin{equation}
\label{JijCPSQUARE}
J_{ij}^{CP} = J_{ij}^0 + t_0^2 \mathcal{E}^2 \left( \frac{1}{\text{U}+\omega} + \frac{1}{\text{U}-\omega} \right)
\end{equation}

\begin{subsection}{Honeycomb lattice}

Now consider a honeycomb lattice structure, so that the $\alpha_{ij}$ values are given by $\alpha_{ij} = \pm|e\bs{R}_{ij}\vec{A}|$ where $\bs{R}_{ij}$ are the displacement vectors in a honeycomb lattice. There are six such vectors:

\begin{align}
\bs{R}_1^\pm &= \pm a\hat{e}_x \\
\bs{R}_2^\pm &= \pm a\frac{1}{2}(\hat{e}_x + \sqrt{3}\hat{e}_y) \\
\bs{R}_3^\pm &= \pm a\frac{1}{2}(\hat{e}_x - \sqrt{3}\hat{e}_y)
\end{align}

These six possible directions will lead to $\alpha_a^\pm = \pm |e\bs{R}_a\vec{A}|$ where $\vec{A}=\frac{iE_0}{\omega\sqrt{1+\lambda^2}}(\hat{e}_x+i\lambda\hat{e}_y)$:

\begin{align}
\alpha_1^\pm &= \pm \frac{eaE_0}{\omega\sqrt{1+\lambda^2}} \\
\alpha_2^\pm &= \alpha_3^\pm = \pm \frac{eaE_0}{2\omega} \sqrt{\frac{1+3\lambda^2}{1+\lambda^2}} = \pm\frac{eaE_0}{2\omega}\sqrt{1+\frac{2\lambda^2}{1+\lambda^2}}
\end{align}

From here we can see that if we take the same approximation as in \ref{Jij2} we will obtain the same form of exchange interaction for circular polarized light ($\lambda=\pm1$) \ref{JijCPSQUARE} with possibly a different lattice constant $a$.

\end{subsection}

\end{section}

\begin{section}{Hubbard model with SOI}
\label{Section3HubbardSOI}
Now, let's investigate the effect of the electric field for the Hubbard model with SOI, i.e. \ref{Hubbard}, in this case the hopping operator gets an extra spin-dependent term:

\begin{equation}
\hat{T} = \sum_{\langle i,j \rangle, \sigma, \sigma'}(\delta_{\sigma, \sigma'} t_0 + \bs{\Delta}_{ij} \bs{\sigma}_{\sigma, \sigma'})\hat{c}_{i \sigma}^\dagger \hat{c}_{j \sigma'}
\end{equation}

The vector $\bs{\Delta}_{ij}$ can describe Rashba or Dresselhaus SOI, $\bs{\Delta}_{ij} = i\Delta_R(R_{ij}^y, -R_{ij}^x, 0)$ for Rashba SOI and $\bs{\Delta}_{ij} = i\Delta_R(R_{ij}^x, -R_{ij}^y, 0)$ for Dresselhaus SOI. 
In this case the hopping amplitudes are:

\begin{equation}
\label{HoppHubbSOI}
t_{ij}^{\sigma \sigma'} = \begin{cases}
	(\delta_{\sigma \sigma'}t_0+\bs{\Delta}_{ij} \bs{\sigma}_{\sigma, \sigma'}) & \text{for } i, j \text{ nearest neighbors} \\
	0 & \text{ otherwise}
\end{cases} \quad
\end{equation}

And as before $\hat{H} = -\hat{T} + \text{U}\hat{D}$. In the presence of an electromagnetic field we can apply the Peierls substitution as we did before, and the effective Hamiltonian will be given by \ref{HeffSimplified}.  inserting the spin operators \ref{SpinOperatorInv1} and \ref{SpinOperatorInv2} we get:

\begin{align}
\hat{H}_{\text{eff}}(t) &= - \frac{1}{2} \sum_{\langle i,j \rangle, \sigma_1, \sigma_2, \sigma_3, \sigma_4}\hat{c}_{i \sigma_1}^\dagger \hat{c}_{j \sigma_2} \hat{c}_{j \sigma_3}^\dagger \hat{c}_{i \sigma_4} t_{ij}^{\sigma_1 \sigma_2} t_{ji}^{\sigma_3 \sigma_4} \mathcal{M}(\alpha_{ij}, \text{U}, \omega, t) \nonumber \\
&= - \frac{1}{2} \sum_{\langle i,j \rangle, \sigma_1, \sigma_2, \sigma_3, \sigma_4} \left( \frac{\delta_{\sigma_1 \sigma_4}}{2} + \bs{S}_i\bs{\sigma}_{\sigma_4 \sigma_1} \right) \left( \frac{\delta_{\sigma_2 \sigma_3}}{2} - \bs{S}_j\bs{\sigma}_{\sigma_2 \sigma_3} \right) \nonumber \\ &(\delta_{\sigma_1 \sigma_2}t_0+\bs{\Delta}_{ij} \bs{\sigma}_{\sigma_1, \sigma_2}) (\delta_{\sigma_3 \sigma_4}t_0+\bs{\Delta}_{ji} \bs{\sigma}_{\sigma_3, \sigma_4}) \mathcal{M}(\alpha_{ij}, \text{U}, \omega, t) \label{HeffHubbSOI1}
\end{align}

Now, since the spin orbit interaction is weaker than the kinetic term, we can neglect the term proportional to $\bs{\Delta}_{ij}^2$ and write:

\begin{equation}
(\delta_{\sigma_1 \sigma_2}t_0+\bs{\Delta}_{ij} \bs{\sigma}_{\sigma_1, \sigma_2}) (\delta_{\sigma_3 \sigma_4}t_0+\bs{\Delta}_{ji} \bs{\sigma}_{\sigma_3, \sigma_4}) \approx \delta_{\sigma_1 \sigma_2}\delta_{\sigma_3 \sigma_4}t_0^2 + t_0\bs{\Delta}_{ij}(\delta_{\sigma_3 \sigma_4} \bs{\sigma}_{\sigma_1 \sigma_2} - \delta_{\sigma_1 \sigma_2} \bs{\sigma}_{\sigma_3 \sigma_4})
\end{equation}

Where we used $\bs{\Delta}_{ji} = - \bs{\Delta}_{ij}$. The term proportional to $t_0^2$ will lead to the modified exchange interaction, and it is exactly the same as in the previous section. The term proportional to $t_0\bs{\Delta}_{ij}$ will lead to the modified DMI interaction.

Now we can rewrite \ref{HeffHubbSOI1} as:

\begin{align*}
\hat{H}_{\text{eff}}(t) &\approx - \frac{1}{2} \sum_{\langle i,j \rangle, \sigma_1, \sigma_2, \sigma_3, \sigma_4} \left( \frac{\delta_{\sigma_1 \sigma_4}}{2} + \bs{S}_i\bs{\sigma}_{\sigma_4 \sigma_1} \right) \left( \frac{\delta_{\sigma_2 \sigma_3}}{2} - \bs{S}_j\bs{\sigma}_{\sigma_2 \sigma_3} \right) \nonumber \\ &\left[ \delta_{\sigma_1 \sigma_2}\delta_{\sigma_3 \sigma_4}t_0^2 + t_0\bs{\Delta}_{ij}(\delta_{\sigma_3 \sigma_4} \bs{\sigma}_{\sigma_1 \sigma_2} - \delta_{\sigma_1 \sigma_2} \bs{\sigma}_{\sigma_3 \sigma_4}) \right] \mathcal{M}(\alpha_{ij}, \text{U}, \omega, t) \\
&= -\frac{t_0^2}{2} \sum_{\langle i,j \rangle \sigma, \sigma'} \left(\frac{1}{2}\delta_{\sigma \sigma'} + \bs{S}_i\bs{\sigma}_{\sigma' \sigma}\right)\left(\frac{1}{2}\delta_{\sigma \sigma'}-\bs{S}_j\bs{\sigma}_{\sigma \sigma'}\right)\mathcal{M}(\alpha_{ij}, \text{U}, \omega, t) - \\
&- \frac{t_0}{2} \sum_{\langle i,j \rangle \sigma_1, \sigma_2, \sigma_3, \sigma_4} \left(\frac{1}{2}\delta_{\sigma_1 \sigma_4} + \bs{S}_i\bs{\sigma}_{\sigma_4 \sigma_1}\right)\left(\frac{1}{2}\delta_{\sigma_2 \sigma_3}-\bs{S}_j\bs{\sigma}_{\sigma_2 \sigma_3}\right) \times \\
&\bs{\Delta}_{ij}(\delta_{\sigma_3,\sigma_4}\bs{\sigma}_{\sigma_1 \sigma_2}-\delta_{\sigma_1,\sigma_2}\bs{\sigma}_{\sigma_3 \sigma_4})\mathcal{M}(\alpha_{ij}, \text{U}, \omega, t) = \\
&= \sum_{\langle i,j \rangle} \left( t_0^2 \bs{S}_i\bs{S}_j + 2it_0\bs{\Delta}_{ij} \bs{S}_i \times \bs{S}_j \right) \mathcal{M}(\alpha_{ij}, \text{U}, \omega, t) =\\
&= \sum_{\langle i,j \rangle} J_{ij}(t)\bs{S}_i\bs{S}_j +\bs{D}_{ij}(t)\bs{S}_i \times \bs{S}_j = \\
&= \hat{H}_J'(t)+\hat{H}_D'(t)
\end{align*}

Where we used relations \ref{SpinRel1} and \ref{SpinRel2}. We have:

\begin{align}
J_{ij}(t) &= t_0^2\mathcal{M}(\alpha_{ij}, \text{U}, \omega, t) \label{JijHSOI} \\
\bs{D}_{ij}(t) &= 2it_0\bs{\Delta}_{ij} \mathcal{M}(\alpha_{ij}, \text{U}, \omega, t) \label{DijHSOI}
\end{align}

Kinetic hopping and the Rashba spin orbit are both NN hopping processes, so they are both renormalized in the same way by the laser field \cite{Stepanov2017}.
For $\bs{D}_{ij}(t)$ this is:

\begin{equation}
\bs{D}_{ij} = 2it_0 \bs{\Delta}_{ij} \sum_{mn}e^{-i(\frac{\pi}{2}+\theta_{ij})m} \left\{ 
    \frac{\mathcal{J}_{n-m}(\alpha_{ij})\mathcal{J}_{n}(\alpha_{ij}) + \mathcal{J}_{n}(\alpha_{ij})\mathcal{J}_{m+n}(\alpha_{ij})}{\text{U}+n\omega} \right\}e^{im\omega t}
\end{equation}

\end{section}

\input{Chapters/KaneMele}

\begin{section}{Other general Hamiltonians}
\label{SectionOtherHams}
\input{Chapters/OtherHams}

\end{section}

\begin{subappendices}
\begin{section}{Peierls Substitution}
\label{AP3A}

In first quantization we can introduce a vector potential $\bs{A}(\bs{r},t)$ by changing the Hamiltonian of the lattice \ref{LaticeHam} to:

\begin{equation}
\label{HamEMF}
  \hat{H}'(\bs{r}) = \frac{(\bs{p}-e\bs{A})^2}{2m} +U(\bs{r})
\end{equation}

Now the Bloch functions defined for \ref{LaticeHam} will not be eigenfunctions of this Hamiltonian. We define a new set of Wannier functions in terms of those defined in \ref{Wannier2}, and obtain the new Bloch functions.

\begin{align}
\psi'_{\bs{R}}(\bs{r}) &= e^{i\frac{e}{\hbar}\int_{\bs{R}}^{\bs{r}} d\bs{r}'\bs{A}(\bs{r}',t)} \psi_{\bs{R}}(\bs{r}) \\
\phi'_{\bs{k}}(\textbf{r}) &= \frac{1}{\sqrt{M}}\sum_{\bs{R}} e^{-i\bs{k}\bs{R}}\psi'_{\bs{R}}(\bs{r}) = \frac{1}{\sqrt{M}}\sum_{\bs{R}} e^{-i\bs{k}\bs{R}} e^{i\frac{e}{\hbar}\int_{\bs{R}}^{\bs{r}} d\bs{r}'\bs{A}(\bs{r}',t)} \psi_{\bs{R}}(\bs{r})
\end{align}

Where we omitted the band number. The action of the new Hamiltonian is greatly simplified:

\begin{align*}
\hat{H}'(\bs{r}) \psi'_{\bs{R}} (\bs{r}) &= \left[ \frac{(\bs{p}-e\bs{A})^2}{2m} +U(\bs{r}) \right] e^{i\frac{e}{\hbar}\int_{\bs{R}}^{\bs{r}} d\bs{r}'\bs{A}(\bs{r}',t)} \psi_{\bs{R}}(\bs{r}) = \\
&=e^{i\frac{e}{\hbar}\int_{\bs{R}}^{\bs{r}} d\bs{r}'\bs{A}(\bs{r}',t)} \left[ \frac{(\bs{p} - i\hbar \bs{\nabla_r}(i\frac{e}{\hbar}\int_{\bs{R}}^{\bs{r}} d\bs{r}'\bs{A}(\bs{r}',t)) -e\bs{A})^2}{2m} +U(\bs{r}) \right] = \\
&= e^{i\frac{e}{\hbar}\int_{\bs{R}}^{\bs{r}} d\bs{r}'\bs{A}(\bs{r}',t)} \left[ \frac{\bs{p}^2}{2m} + U(\bs{r}) \right] \psi_{\bs{R}}(\bs{r}) = e^{i\frac{e}{\hbar}\int_{\bs{R}}^{\bs{r}} d\bs{r}'\bs{A}(\bs{r}',t)} \hat{H}(\bs{r}) \psi_{\bs{R}}(\bs{r})
\end{align*}

And so, the matrix elements as in the case without field except for a phase:

\begin{equation}
\label{PeierlsMtxElem}
t_{ij}(t) = \bra{\psi'_{\bs{R_i}}} \hat{H}'(\bs{r}) \ket{\psi'_{\bs{R_j}}} = e^{i\frac{e}{\hbar}\int_{\bs{R}_j}^{\bs{R}_i} d\bs{r}'\bs{A}(\bs{r}',t)} \bra{\psi_{\bs{R_i}}} \hat{H}(\bs{r}) \ket{\psi_{\bs{R_j}}}
\end{equation}

When the field can be approximated as constant along the lattice we have: $e^{i\frac{e}{\hbar}\int_{\bs{R}_j}^{\bs{R}_i} d\bs{r}'\bs{A}(\bs{r}',t)} = e^{i\frac{e}{\hbar} (\bs{R}_i-\bs{R}_j) \bs{A}(t)}$.

Notice that this substitution also applies if we add a Rashba Spin-Orbit interaction term to \ref{HamEMF}, i.e., if $\hat{H}'(\bs{r}) = \frac{(\bs{p}-e\bs{A})^2}{2m} +U(\bs{r}) + \alpha_R \hat{z}(\hat{\bs{\sigma}} \times (\bs{p}-e\bs{A}))$, then

\begin{align*}
  \alpha_R \hat{z}(\hat{\bs{\sigma}} \times (\bs{p}-e\bs{A})) \psi'_{\bs{R}} &= \alpha_R(\hat{\sigma_x}(\hat{p}_y-e\bs{A}) - \hat{\sigma_y}(\hat{p}_x-e\bs{A})) e^{i\frac{e}{\hbar}\int_{\bs{R}}^{\bs{r}}d\bs{r}'\bs{A}(\bs{r}',t)} \psi_{\bs{R}} = \\
  &= e^{i\frac{e}{\hbar}\int_{\bs{R}}^{\bs{r}}d\bs{r}'\bs{A}(\bs{r}',t)} \alpha_R \hat{z}\hat{\bs{\sigma}} \times \bs{p} \psi_{\bs{R}}
\end{align*}

Therefore relation \ref{PeierlsMtxElem} still holds.

\end{section}

\begin{section}{Identities}
\label{AP3B}
Let $Z(t)$ be an operator dependent on the parameter $t$. Then:

\begin{align*}
d_t e^{Z} &= d_t \sum_{n=0}^\infty \frac{1}{n!} Z^n = \sum_{n=0}^\infty \sum_{m=0}^{n-1} \frac{1}{n!} Z^m \frac{\partial Z}{\partial t} Z^{n-m-1} = \\
&= \sum_{m=0}^\infty \sum_{p=0}^\infty \frac{1}{(m+p+1)!}Z^m \frac{\partial Z}{\partial t} Z^p = \sum_{m=0}^\infty \sum_{p=0}^\infty \frac{m!p!}{(m+p+1)!} \frac{Z^m}{m!} \frac{\partial Z}{\partial t} \frac{Z^p}{p!} =\\
&= \sum_{m=0}^\infty \sum_{p=0}^\infty \int_0^1 dx x^m (1-x)^p \frac{Z^m}{m!} \frac{\partial Z}{\partial t} \frac{Z^p}{p!} = \int_0^1 dx e^{Zx} \frac{\partial Z}{\partial t} e^{Z(1-x)} = \\
&= \int_0^1 dx \sum_{n=0}^\infty \frac{1}{n!} \left[xZ, \dots, \left[xZ, \frac{\partial Z}{\partial t}\right] \dots \right] e^Z = \sum_{n=0}^\infty \frac{1}{(n+1)!} \left[Z, \dots, \left[Z, \frac{\partial Z}{\partial t}\right] \dots \right] e^Z
\end{align*}

Where we used the Beta function $\int_0^1 dx x^m (1-x)^p = \frac{m!p!}{(m+p+1)!}$ and we also used the Baker–Campbell–Hausdorff expansion for $e^{xZ}\frac{\partial Z}{\partial t}e^{-xZ}$ and $\int_0^1 x^n = \frac{1}{n+1}$. Therefore we proved:

\begin{equation}
\label{SneidersID}
d_t e^{Z} = \sum_{n=0}^\infty \frac{1}{(n+1)!} \left[Z, \dots, \left[Z, \frac{\partial Z}{\partial t}\right] \dots \right] e^Z = \sum_{n=0}^\infty \frac{1}{(n+1)!} \text{ad}_Z ^n \left(\frac{\partial Z}{\partial t}\right) e^Z
\end{equation}

Where $\text{ad}_A(B) = [A,B]$.

\end{section}

\end{subappendices}