### 2019-02-25 Appendices/AppendixE.tex
Fix references

\chapter{Floquet Theory}
\label{APE}
In this appendix we will state and proof the Floquet theorem and employ the Floquet formalism to derive a perturbative analysis of high frequency time periodic Hamiltonians. This formalism been widely used recently in many-body driven systems (\cite{Desbuquois2017}, \cite{Bordia2017}, \cite{Gorg2018}).