### 2019-04-01 Chapters/Chapter5.tex
more spin waves

\chapter{Spin wave theory of antiferromagnets}
\label{SpinWave}

\section{Heisenberg antiferromagnet}

We will start with the Heisenberg antiferromagnet:

\begin{equation}
\hat{H} = J \sum_{\langle i j \rangle} \bs{S}_i \bs{S}_j
\end{equation}

In order to obtain the magnon Hamiltonian for this model we introduce the antiferromagnet Holstein-Primakoff transformation:

\begin{align}
S^+_{Ai} &= \sqrt{2S-\hat{a}^\dagger_i \hat{a}_i} \hat{a}_i \\
S^-_{Ai} &= \hat{a}^\dagger_i\sqrt{2S-\hat{a}^\dagger_i \hat{a}_i}  \\
S^z_{Ai} &= S - \hat{a}^\dagger_i \hat{a}_i \\
S^+_{Bj} &=\hat{b}^\dagger_j\sqrt{2S-\hat{b}^\dagger_j \hat{b}_j} \\
S^-_{Bj} &= \sqrt{2S-\hat{b}^\dagger_j \hat{b}_j} \hat{b}_j \\
S^z_{Bj} &= -S + \hat{b}^\dagger_j \hat{b}_j
\end{align}

Where $A$ and $B$ denote two Ne\'{e}l sublattices and $\hat{a}_i$, $\hat{b}_i$ are bosonic annihilation operators. In the large-S limit we approximate $ \sqrt{2S-\hat{a}^\dagger_i \hat{a}_i} \approx \sqrt{2S}$. Then, the exchange term becomes:

\begin{align*}
\bs{S}_i\bs{S}_j = S_i^zS_j^z + \frac{1}{2}\left(S_i^+S_j^- + S_i^-S_j^+ \right) = S \left( \hat{a}^\dagger_i \hat{a}_i + \hat{b}^\dagger_j \hat{b}_j + \hat{a}_i\hat{b}_j + \hat{a}^\dagger_i\hat{b}^\dagger_j \right) - \hat{a}^\dagger_i \hat{a}_i\hat{b}^\dagger_j \hat{b}_j - S^2
\end{align*}

Now the $S^2$ term is a constant that lowers the energy of the antiferromagnet, therefore is shall not be taken into account to compute the magnon dispersion. The term $\hat{a}^\dagger_i \hat{a}_i\hat{b}^\dagger_j \hat{b}_j$ is zeroth order in $S$ and therefore neglected. We can rewrite the sum over NN as:

\begin{equation}
\sum_{\langle i j \rangle} = 2\sum_{i \in A \delta}
\end{equation}

Where $\delta$ are the NN vectors. Using this, and the Fourier relations:

\begin{align}
\hat{a}_i^\dagger &= \frac{1}{\sqrt{N}} \sum_{\bs{k}} e^{i \bs{k} \cdot \bs{r}_i} \hat{a}_{\bs{k}}^\dagger \\
\hat{b}_j^\dagger &= \frac{1}{\sqrt{N}} \sum_{\bs{k}} e^{i \bs{k} \cdot \bs{r}_j} \hat{b}_{\bs{k}}^\dagger
\end{align}

Where the sum is over $\bs{k}$ in the first Brillouin zone. Using these relations, together with $\sum{i} e^{i{\bs{k}-\bs{k}'}\cdot\bs{r}_i} = \delta_{\bs{k}, \bs{k}'}$ we find:

\begin{equation}
\hat{H} = 2JS \sum_{\bs{k} \bs{\delta}} \left( \hat{a}^\dagger_{\bs{k}}\hat{a}_{\bs{k}} + \hat{b}^\dagger_{\bs{k}}\hat{b}_{\bs{k}} + e^{i\bs{k}\bs{\delta}}\hat{a}_{\bs{k}}\hat{b}_{\bs{-k}} + e^{-i \bs{k}\cdot\bs{\delta}}  \hat{a}^\dagger_{\bs{k}}\hat{b}^\dagger_{-\bs{k}} \right)
\end{equation}

At this point it is common to define $\gamma_{\bs{k}} = \sum_\delta e^{i \bs{k} \cdot \bs{\delta}}$, then denoting $z = \sum_\delta$ we get:

\begin{equation}
\label{MagnonicH}
\hat{H} = 2JS \sum_{\bs{k}} \left\{ z \left( \hat{a}^\dagger_{\bs{k}}\hat{a}_{\bs{k}} + \hat{b}^\dagger_{-\bs{k}}\hat{b}_{-\bs{k}} \right) + \gamma_{\bs{k}}\hat{a}_{\bs{k}}\hat{b}_{\bs{-k}} + \gamma_{-\bs{k}}\hat{a}^\dagger_{\bs{k}}\hat{b}^\dagger_{-\bs{k}} \right\}
\end{equation}

Which can be written in matrix form as:

\begin{equation}
\hat{H}_{\bs{k}} = 2JS\begin{pmatrix} 
z & \gamma_{\bs{k}} \\
\gamma_{-\bs{k}} & z
\end{pmatrix}
\end{equation}

Acting on the spinor $\Psi_{\bs{k}} = \left(\hat{a}_{\bs{k}},\hat{b}_{-\bs{k}}^\dagger \right)^T$. That is, $\hat{H} = \sum_{\bs{k}} \Psi_{\bs{k}} \hat{H}_{\bs{k}} \Psi^\dagger_{\bs{k}}$. In order to diagonalize this Hamiltonian we introduce the Bogoliubov transformation:

\begin{equation}
\begin{pmatrix}
\hat{\alpha}_{\bs{k}} \\
\hat{\beta}_{\bs{k}}^\dagger
\end{pmatrix} = 
\begin{pmatrix}
u_{\bs{k}} & -v_{\bs{k}} \\
-v_{\bs{k}} & u_{\bs{k}}
\end{pmatrix}
\begin{pmatrix}
\hat{a}_{\bs{k}} \\
\hat{b}_{-\bs{k}}^\dagger
\end{pmatrix}
\end{equation}

Where $u_{\bs{k}}$ and $v_{\bs{k}}$ are real functions satisfying $u_{\bs{k}}^2-v_{\bs{k}}^2 = 1$ in order to perserve the bosonic commutation relations. The inverse transformation is given by:

\begin{equation}
\begin{pmatrix}
\hat{a}_{\bs{k}} \\
\hat{b}_{-\bs{k}}^\dagger
\end{pmatrix} = 
\begin{pmatrix}
u_{\bs{k}} & v_{\bs{k}} \\
v_{\bs{k}} & u_{\bs{k}}
\end{pmatrix}
\begin{pmatrix}
\hat{\alpha}_{\bs{k}} \\
\hat{\beta}_{\bs{k}}^\dagger
\end{pmatrix}
\end{equation}

Inserting this in \ref{MagnonicH} and ignoring constant terms we get:

\begin{align*}
\hat{H} &= 2JS \sum_{\bs{k}} \left\{ \left( \hat{\alpha}_{\bs{k}}^\dagger \hat{\alpha}_{\bs{k}} +\hat{\beta}_{\bs{k}}^\dagger \hat{\beta}_{\bs{k}} \right) \left( z(u_{\bs{k}}^2+v_{\bs{k}}^2) +u_{\bs{k}}v_{\bs{k}}( \gamma_{\bs{k}}+ \gamma_{-\bs{k}}) \right) + \right. \\
&+ \left.  \hat{\alpha}^\dagger_{\bs{k}} \hat{\beta}^\dagger_{\bs{k}} \left( 2zu_{\bs{k}}v_{\bs{k}} + \gamma_{\bs{k}}v_{\bs{k}}^2+\gamma_{-\bs{k}}u_{\bs{k}}^2 \right) + \hat{\alpha}_{\bs{k}}\hat{\beta}_{\bs{k}} \left( 2zu_{\bs{k}}v_{\bs{k}} + \gamma_{\bs{k}}u_{\bs{k}}^2+\gamma_{-\bs{k}}v_{\bs{k}}^2 \right) \right\}
\end{align*}

($\gamma_{\bs{k}} = \gamma_{-\bs{k}}$)Therefore, in order to erase the non diagonal terms we impose $2zu_{\bs{k}}v_{\bs{k}} + \gamma_{\bs{k}}(u_{\bs{k}}^2+v_{\bs{k}}^2) = 0$, which together with $u_{\bs{k}}^2-v_{\bs{k}}^2 = 1$ determines $u_{\bs{k}}$ and $v_{\bs{k}}$:

\begin{align*}
u_{\bs{k}}^2 &= \frac{1}{2}\left( \frac{1}{\sqrt{1-(\gamma_{\bs{k}}/z)^2}}+1\right) \\
v_{\bs{k}}^2 &= \frac{1}{2}\left( \frac{1}{\sqrt{1-(\gamma_{\bs{k}}/z)^2}}-1\right)
\end{align*}

And so, the we get the dispersion relation:

\begin{align}
\hat{H} &= \sum_{\bs{k}} \omega_{\bs{k}} \left( \hat{\alpha}_{\bs{k}}^\dagger \hat{\alpha}_{\bs{k}} +\hat{\beta}_{\bs{k}}^\dagger \hat{\beta}_{\bs{k}} \right) \\
\omega_{\bs{k}} &= 2JSz\sqrt{1-(\gamma_{\bs{k}}/z)^2}
\end{align}

\section{Adding DMI}

Let's now consider

\begin{equation}
\hat{H} = J \sum_{\langle i j \rangle} \bs{S}_i \bs{S}_j + D \sum_{\langle \langle i j \rangle \rangle} \nu_{ij}\hat{e}_z \bs{S}_i \times \bs{S}_j
\end{equation}

The NNN DMI term will couple spins within the same sublattice. Using the HP transformation:

\begin{equation}
\hat{e}_z \bs{S}_i \times \bs{S}_j = \frac{i}{2}(S_i^+S_j^--S_i^-S_j^+) = \begin{cases}
             iS(\hat{a}_j^\dagger\hat{a}_i-\hat{a}_i^\dagger\hat{a}_j),  & \text{for } (i,j) \in A \\
             iS(\hat{b}_i^\dagger\hat{b}_j-\hat{b}_j^\dagger\hat{b}_i),  & \text{for } (i,j) \in B
       \end{cases} \quad
\end{equation}

Now we will make the sum over NNN in the following way:

\begin{equation}
\sum_{\langle \langle i j \rangle \rangle} = \sum_{i\in A \bs{\delta}^{NNN}} +  \sum_{i\in B \bs{\delta}^{NNN}}
\end{equation}

Where $\bs{\delta}^{NNN}$ are the NNN vectors. Following the same procedure as before we get and additional magnonic term:

\begin{equation}
D \sum_{\langle \langle i j \rangle \rangle} \nu_{ij}\hat{e}_z \bs{S}_i \times \bs{S}_j = 2JS\sum_{\bs{k}} \left\{ \Delta_{\bs{k}}\hat{a}^\dagger_{\bs{k}}\hat{a}_{\bs{k}} - \Delta_{-\bs{k}}\hat{b}^\dagger_{-\bs{k}}\hat{b}_{-\bs{k}} \right\}
\end{equation}

Where $\Delta_{\bs{k}} = \frac{D}{2J}\sum_{\bs{\delta}^{NNN}} \nu_{i,i+\delta} (e^{i\bs{k}\cdot\bs{\delta}} -e^{-i\bs{k}\cdot\bs{\delta}})$. The full magnonic Hamiltonian now is:

\begin{equation}
\label{MagnonicH2}
\hat{H} = 2JS \sum_{\bs{k}} \left\{ \hat{a}^\dagger_{\bs{k}}\hat{a}_{\bs{k}}(z+\Delta_{\bs{k}}) + \hat{b}^\dagger_{-\bs{k}}\hat{b}_{-\bs{k}}(z-\Delta_{-\bs{k}}) + \gamma_{\bs{k}}\hat{a}_{\bs{k}}\hat{b}_{\bs{-k}} + \gamma_{-\bs{k}}\hat{a}^\dagger_{\bs{k}}\hat{b}^\dagger_{-\bs{k}} \right\}
\end{equation}

Or:

\begin{equation}
\hat{H}_{\bs{k}} = 2JS\begin{pmatrix} 
z + \Delta_{\bs{k}}& \gamma_{\bs{k}} \\
\gamma_{-\bs{k}} & z - \Delta_{-\bs{k}}
\end{pmatrix}
\end{equation}

\section{NNN exchange}

Let's now introduce a term:

\begin{equation}
\sum_{\langle \langle i,j \rangle \rangle} \left\{ \Gamma_{xy}(S_i^xS_j^x + S_i^yS_j^y) + \Gamma_zS_i^zS_j^z\right\}
\end{equation}
In terms of bosonic operators we have:
\begin{equation}
\Gamma_{xy}(S_i^xS_j^x + S_i^yS_j^y) + \Gamma_zS_i^zS_j^z = \begin{cases}
             S\Gamma_{xy}(\hat{a}_i^\dagger\hat{a}_j+\hat{a}_j^\dagger\hat{a}_i) - S\Gamma_z(\hat{a}_i^\dagger\hat{a}_i+\hat{a}_j^\dagger\hat{a}_j),  & \text{for } (i,j) \in A \\
             S\Gamma_{xy}(\hat{b}_i^\dagger\hat{b}_j+\hat{b}_j^\dagger\hat{b}_i) - S\Gamma_z(\hat{b}_i^\dagger\hat{b}_i+\hat{b}_j^\dagger\hat{b}_j),  & \text{for } (i,j) \in B
       \end{cases} \quad
\end{equation}
Fourier transform leads to:

\begin{align*}
&\sum_{\langle \langle i,j \rangle \rangle} \left\{ \Gamma_{xy}(S_i^xS_j^x + S_i^yS_j^y) + \Gamma_zS_i^zS_j^z\right\}
= S\sum_{\bs{k}} (\Gamma_{xy}\Delta'_{\bs{k}}-4z\Gamma_z) (\hat{a}^\dagger_{\bs{k}}\hat{a}_{\bs{k}}+\hat{b}^\dagger_{-\bs{k}}\hat{b}_{-\bs{k}}) = \\
&=2JS\sum_{\bs{k}} \tilde{\Gamma}_{\bs{k}}(\hat{a}^\dagger_{\bs{k}}\hat{a}_{\bs{k}}+\hat{b}^\dagger_{-\bs{k}}\hat{b}_{-\bs{k}})
\end{align*}

Where $\Delta'_{\bs{k}} = \sum_{\delta^{NNN}} \left( e^{i\bs{k}\cdot \bs{\delta}^{NNN}} +  e^{-i\bs{k}\cdot \bs{\delta}^{NNN}}\right)$ and $\tilde{\Gamma}_{\bs{k}} = \frac{1}{2J}\left( \Gamma_{xy}\Delta'_{\bs{k}}-4z\Gamma_z \right)$. Therefore, now:

\begin{equation}
\hat{H}_{\bs{k}} = 2JS\begin{pmatrix} 
z + \Delta_{\bs{k}} + \tilde{\Gamma}_{\bs{k}} & \gamma_{\bs{k}} \\
\gamma_{-\bs{k}} & z - \Delta_{-\bs{k}} + \tilde{\Gamma}_{\bs{k}}
\end{pmatrix}
\end{equation}