### 2019-04-04 Chapters/Chapter5.tex
Add dispersion relations

\chapter{Spin wave theory of antiferromagnets}
\label{SpinWave}

\section{Heisenberg antiferromagnet}

We will start with the Heisenberg antiferromagnet:

\begin{equation}
\hat{H} = J \sum_{\langle i j \rangle} \bs{S}_i \bs{S}_j
\end{equation}

In order to obtain the magnon Hamiltonian for this model we introduce the antiferromagnet Holstein-Primakoff transformation:

\begin{align}
S^+_{Ai} &= \sqrt{2S-\hat{a}^\dagger_i \hat{a}_i} \hat{a}_i \\
S^-_{Ai} &= \hat{a}^\dagger_i\sqrt{2S-\hat{a}^\dagger_i \hat{a}_i}  \\
S^z_{Ai} &= S - \hat{a}^\dagger_i \hat{a}_i \\
S^+_{Bj} &=\hat{b}^\dagger_j\sqrt{2S-\hat{b}^\dagger_j \hat{b}_j} \\
S^-_{Bj} &= \sqrt{2S-\hat{b}^\dagger_j \hat{b}_j} \hat{b}_j \\
S^z_{Bj} &= -S + \hat{b}^\dagger_j \hat{b}_j
\end{align}

Where $A$ and $B$ denote two Ne\'{e}l sublattices and $\hat{a}_i$, $\hat{b}_i$ are bosonic annihilation operators. In the large-S limit we approximate $ \sqrt{2S-\hat{a}^\dagger_i \hat{a}_i} \approx \sqrt{2S}$. Then, the exchange term becomes:

\begin{align*}
\bs{S}_i\bs{S}_j = S_i^zS_j^z + \frac{1}{2}\left(S_i^+S_j^- + S_i^-S_j^+ \right) = S \left( \hat{a}^\dagger_i \hat{a}_i + \hat{b}^\dagger_j \hat{b}_j + \hat{a}_i\hat{b}_j + \hat{a}^\dagger_i\hat{b}^\dagger_j \right) - \hat{a}^\dagger_i \hat{a}_i\hat{b}^\dagger_j \hat{b}_j - S^2
\end{align*}

Now the $S^2$ term is a constant that lowers the energy of the antiferromagnet, therefore is shall not be taken into account to compute the magnon dispersion. The term $\hat{a}^\dagger_i \hat{a}_i\hat{b}^\dagger_j \hat{b}_j$ is zeroth order in $S$ and therefore neglected. We can rewrite the sum over NN as:

\begin{equation}
\sum_{\langle i j \rangle} = 2\sum_{i \in A \delta}
\end{equation}

Where $\delta$ are the NN vectors. Using this, and the Fourier relations:

\begin{align}
\hat{a}_i^\dagger &= \frac{1}{\sqrt{N}} \sum_{\bs{k}} e^{i \bs{k} \cdot \bs{r}_i} \hat{a}_{\bs{k}}^\dagger \\
\hat{b}_j^\dagger &= \frac{1}{\sqrt{N}} \sum_{\bs{k}} e^{i \bs{k} \cdot \bs{r}_j} \hat{b}_{\bs{k}}^\dagger
\end{align}

Where the sum is over $\bs{k}$ in the first Brillouin zone. Using these relations, together with $\sum{i} e^{i{\bs{k}-\bs{k}'}\cdot\bs{r}_i} = \delta_{\bs{k}, \bs{k}'}$ we find:

\begin{equation}
\hat{H} = 2JS \sum_{\bs{k} \bs{\delta}} \left( \hat{a}^\dagger_{\bs{k}}\hat{a}_{\bs{k}} + \hat{b}^\dagger_{\bs{k}}\hat{b}_{\bs{k}} + e^{i\bs{k}\bs{\delta}}\hat{a}_{\bs{k}}\hat{b}_{\bs{-k}} + e^{-i \bs{k}\cdot\bs{\delta}}  \hat{a}^\dagger_{\bs{k}}\hat{b}^\dagger_{-\bs{k}} \right)
\end{equation}

At this point it is common to define $\gamma_{\bs{k}} = \sum_\delta e^{i \bs{k} \cdot \bs{\delta}}$, then denoting $z = \sum_\delta$ we get:

\begin{equation}
\label{MagnonicH}
\hat{H} = 2JS \sum_{\bs{k}} \left\{ z \left( \hat{a}^\dagger_{\bs{k}}\hat{a}_{\bs{k}} + \hat{b}^\dagger_{-\bs{k}}\hat{b}_{-\bs{k}} \right) + \gamma_{\bs{k}}\hat{a}_{\bs{k}}\hat{b}_{\bs{-k}} + \gamma_{-\bs{k}}\hat{a}^\dagger_{\bs{k}}\hat{b}^\dagger_{-\bs{k}} \right\}
\end{equation}

Which can be written in matrix form as:

\begin{equation}
\hat{H}_{\bs{k}} = 2JS\begin{pmatrix} 
z & \gamma_{\bs{k}} \\
\gamma_{-\bs{k}} & z
\end{pmatrix}
\end{equation}

Acting on the spinor $\Psi_{\bs{k}} = \left(\hat{a}_{\bs{k}},\hat{b}_{-\bs{k}}^\dagger \right)^T$. That is, $\hat{H} = \sum_{\bs{k}} \Psi_{\bs{k}} \hat{H}_{\bs{k}} \Psi^\dagger_{\bs{k}}$. In order to diagonalize this Hamiltonian we introduce the Bogoliubov transformation:

\begin{equation}
\label{Bogoliubov}
\begin{pmatrix}
\hat{\alpha}_{\bs{k}} \\
\hat{\beta}_{\bs{k}}
\end{pmatrix} = 
\begin{pmatrix}
u_{\bs{k}} & v_{\bs{k}}^* \\
v_{\bs{k}} & u_{\bs{k}}^*
\end{pmatrix}
\begin{pmatrix}
\hat{a}_{\bs{k}} \\
\hat{b}_{-\bs{k}}^\dagger
\end{pmatrix}
\end{equation}

Where $u_{\bs{k}}$ and $v_{\bs{k}}$ are complex functions satisfying $\abs{u_{\bs{k}}}^2-\abs{v_{\bs{k}}}^2 = 1$ in order to perserve the bosonic commutation relations. The inverse transformation is given by:

\begin{equation}
\label{BogoliubovInv}
\begin{pmatrix}
\hat{a}_{\bs{k}} \\
\hat{b}_{-\bs{k}}^\dagger
\end{pmatrix} = 
\begin{pmatrix}
u_{\bs{k}}^* & -v_{\bs{k}}^* \\
-v_{\bs{k}} & u_{\bs{k}}
\end{pmatrix}
\begin{pmatrix}
\hat{\alpha}_{\bs{k}} \\
\hat{\beta}_{\bs{k}}^\dagger
\end{pmatrix}
\end{equation}

Inserting this in \ref{MagnonicH} and ignoring constant terms we get:

\begin{align}
\hat{H} &= 2JS \sum_{\bs{k}} \left\{ \left( \hat{\alpha}_{\bs{k}}^\dagger \hat{\alpha}_{\bs{k}} +\hat{\beta}_{\bs{k}}^\dagger \hat{\beta}_{\bs{k}} \right) \left( z(\abs{u_{\bs{k}}}^2+\abs{v_{\bs{k}}}^2) -\gamma_{\bs{k}}u_{\bs{k}}^*v_{\bs{k}}^* - \gamma_{-\bs{k}}u_{\bs{k}}v_{\bs{k}} \right) + \right. \nonumber \\
&+ \left.  \hat{\alpha}^\dagger_{\bs{k}} \hat{\beta}_{\bs{k}} \left( -2zu_{\bs{k}}v_{\bs{k}}^* + \gamma_{\bs{k}}(v_{\bs{k}}^*)^2+\gamma_{-\bs{k}}u_{\bs{k}}^2 \right) + \hat{\beta}_{\bs{k}}^\dagger\hat{\alpha}_{\bs{k}} \left( -2zu_{\bs{k}}^*v_{\bs{k}} + \gamma_{\bs{k}}(u_{\bs{k}}^*)^2+\gamma_{-\bs{k}}v_{\bs{k}}^2 \right) \right\} \label{BeforeBogo}
\end{align}

Therefore, in order to erase the non diagonal terms we impose:
\begin{equation}
\label{condition}
-2zu_{\bs{k}}v_{\bs{k}}^* + \gamma_{\bs{k}}(v_{\bs{k}}^*)^2+\gamma_{-\bs{k}}u_{\bs{k}}^2 = 0
\end{equation}
Notice that for the $\hat{\beta}_{\bs{k}}^\dagger\hat{\alpha}_{\bs{k}}$ term $-2zu_{\bs{k}}^*v_{\bs{k}} + \gamma_{\bs{k}}(u_{\bs{k}}^*)^2+\gamma_{-\bs{k}}v_{\bs{k}}^2=0$ is the same equation. This, together with $\abs{u_{\bs{k}}}^2-\abs{v_{\bs{k}}}^2 = 1$ determines $u_{\bs{k}}$ and $v_{\bs{k}}$. To see this, first notice that an overall phase in both $u_{\bs{k}}$ and $v_{\bs{k}}$ leads to the same physical state. We can use this fact to impose that $v_{\bs{k}}$ is real, and we can use $\abs{u_{\bs{k}}}^2-\abs{v_{\bs{k}}}^2 = 1$ to parametrize $u_{\bs{k}}$ and $v_{\bs{k}}$ in the following way:

\begin{align*}
u_{\bs{k}} &= \cosh(\frac{\theta_{\bs{k}}}{2})e^{i \xi_{\bs{k}}} \\
v_{\bs{k}} &= \sinh(\frac{\theta_{\bs{k}}}{2})
\end{align*}

Where $\theta_{\bs{k}}$ and $\xi_{\bs{k}}$ are real functions to be determined. Introducing these relations in \ref{condition}:

\begin{equation}
- z\sinh(\theta_{\bs{k}}) +\gamma_{\bs{k}}e^{-i \xi_{\bs{k}}}\frac{ \cosh(\theta_{\bs{k}})-1}{2} +\gamma_{-\bs{k}}e^{i\xi_{\bs{k}}}\frac{ \cosh(\theta_{\bs{k}})+1}{2} = 0
\end{equation}

By taking $\xi_{\bs{k}}$ to be the phase of $\gamma_{\bs{k}}$, i.e. $\gamma_{\bs{k}} = \abs{\gamma_{\bs{k}}} e^{i\xi_{\bs{k}}}$ we get a real equation on $\theta_{\bs{k}}$. Altogether we find:

\begin{align}
\gamma_{\bs{k}} &= \abs{\gamma_{\bs{k}}} e^{i\xi_{\bs{k}}} \\
\tanh(\theta_{\bs{k}}) &= \frac{\abs{\gamma_{\bs{k}}}}{z}
\end{align}

This erases the non diagonal terms in \ref{BeforeBogo} leading to a diagonal term:
\begin{equation}
z\cosh(\theta_{\bs{k}})-\abs{\gamma_{\bs{k}}}\sinh(\theta_{\bs{k}}) = \sqrt{z^2-\abs{\gamma_{\bs{k}}}^2}
\end{equation}

Then, the dispersion relation is:

\begin{align}
\hat{H} &= \sum_{\bs{k}}\left( \hat{\alpha}_{\bs{k}}^\dagger \hat{\alpha}_{\bs{k}} +\hat{\beta}_{\bs{k}}^\dagger \hat{\beta}_{\bs{k}} \right) \epsilon_{\bs{k}} \\
\epsilon_{\bs{k}} &= 2JS \sqrt{z^2-\abs{\gamma_{\bs{k}}}^2}
\end{align}

\section{Adding DMI}

Let's now consider

\begin{equation}
\hat{H} = J \sum_{\langle i j \rangle} \bs{S}_i \bs{S}_j + D \sum_{\langle \langle i j \rangle \rangle} \nu_{ij}\hat{e}_z \bs{S}_i \times \bs{S}_j
\end{equation}

The NNN DMI term will couple spins within the same sublattice. Using the HP transformation:

\begin{equation}
\hat{e}_z \bs{S}_i \times \bs{S}_j = \frac{i}{2}(S_i^+S_j^--S_i^-S_j^+) = \begin{cases}
             iS(\hat{a}_j^\dagger\hat{a}_i-\hat{a}_i^\dagger\hat{a}_j),  & \text{for } (i,j) \in A \\
             iS(\hat{b}_i^\dagger\hat{b}_j-\hat{b}_j^\dagger\hat{b}_i),  & \text{for } (i,j) \in B
       \end{cases} \quad
\end{equation}

Now we will make the sum over NNN in the following way:

\begin{equation}
\sum_{\langle \langle i j \rangle \rangle} = \sum_{i\in A \bs{\delta}^{NNN}} +  \sum_{i\in B \bs{\delta}^{NNN}}
\end{equation}

Where $\bs{\delta}^{NNN}$ are the NNN vectors. Following the same procedure as before we get and additional magnonic term:

\begin{equation}
D \sum_{\langle \langle i j \rangle \rangle} \nu_{ij}\hat{e}_z \bs{S}_i \times \bs{S}_j = 2JS\sum_{\bs{k}} \left\{ \Delta_{\bs{k}}\hat{a}^\dagger_{\bs{k}}\hat{a}_{\bs{k}} - \Delta_{-\bs{k}}\hat{b}^\dagger_{-\bs{k}}\hat{b}_{-\bs{k}} \right\}
\end{equation}

Where $\Delta_{\bs{k}} = \frac{D}{2J}\sum_{\bs{\delta}^{NNN}} \nu_{i,i+\delta} (e^{i\bs{k}\cdot\bs{\delta}} -e^{-i\bs{k}\cdot\bs{\delta}})$. The full magnonic Hamiltonian now is:

\begin{equation}
\label{MagnonicH2}
\hat{H} = 2JS \sum_{\bs{k}} \left\{ \hat{a}^\dagger_{\bs{k}}\hat{a}_{\bs{k}}(z+\Delta_{\bs{k}}) + \hat{b}^\dagger_{-\bs{k}}\hat{b}_{-\bs{k}}(z-\Delta_{-\bs{k}}) + \gamma_{\bs{k}}\hat{a}_{\bs{k}}\hat{b}_{\bs{-k}} + \gamma_{-\bs{k}}\hat{a}^\dagger_{\bs{k}}\hat{b}^\dagger_{-\bs{k}} \right\}
\end{equation}

Or:

\begin{equation}
\hat{H}_{\bs{k}} = 2JS\begin{pmatrix} 
z + \Delta_{\bs{k}}& \gamma_{\bs{k}} \\
\gamma_{-\bs{k}} & z - \Delta_{-\bs{k}}
\end{pmatrix}
\end{equation}

In this case, applying the transformation \ref{Bogoliubov} leads to the same condition for the coefficients  $u_{\bs{k}}$ and $v_{\bs{k}}$, whereas the energy gets lifted:

\begin{align}
\hat{H} &= \sum_{\bs{k}} \left( \hat{\alpha}_{\bs{k}}^\dagger \hat{\alpha}_{\bs{k}} \epsilon_{\bs{k}}^+ +\hat{\beta}_{\bs{k}}^\dagger \hat{\beta}_{\bs{k}}  \epsilon_{\bs{k}}^- \right) \label{DispWithDMI} \\
\epsilon_{\bs{k}}^\pm &= 2JS \left( \pm \Delta_{\bs{k}} + \sqrt{z^2-\abs{\gamma_{\bs{k}}}^2} \right)
\end{align}

\section{NNN exchange}

Let's now introduce a term:

\begin{equation}
\sum_{\langle \langle i,j \rangle \rangle} \left\{ \Gamma_{xy}(S_i^xS_j^x + S_i^yS_j^y) + \Gamma_zS_i^zS_j^z\right\}
\end{equation}
In terms of bosonic operators we have:
\begin{equation}
\Gamma_{xy}(S_i^xS_j^x + S_i^yS_j^y) + \Gamma_zS_i^zS_j^z = \begin{cases}
             S\Gamma_{xy}(\hat{a}_i^\dagger\hat{a}_j+\hat{a}_j^\dagger\hat{a}_i) - S\Gamma_z(\hat{a}_i^\dagger\hat{a}_i+\hat{a}_j^\dagger\hat{a}_j),  & \text{for } (i,j) \in A \\
             S\Gamma_{xy}(\hat{b}_i^\dagger\hat{b}_j+\hat{b}_j^\dagger\hat{b}_i) - S\Gamma_z(\hat{b}_i^\dagger\hat{b}_i+\hat{b}_j^\dagger\hat{b}_j),  & \text{for } (i,j) \in B
       \end{cases} \quad
\end{equation}
Fourier transform leads to:

\begin{align*}
&\sum_{\langle \langle i,j \rangle \rangle} \left\{ \Gamma_{xy}(S_i^xS_j^x + S_i^yS_j^y) + \Gamma_zS_i^zS_j^z\right\}
= S\sum_{\bs{k}} (\Gamma_{xy}\Delta'_{\bs{k}}-4z\Gamma_z) (\hat{a}^\dagger_{\bs{k}}\hat{a}_{\bs{k}}+\hat{b}^\dagger_{-\bs{k}}\hat{b}_{-\bs{k}}) = \\
&=2JS\sum_{\bs{k}} \tilde{\Gamma}_{\bs{k}}(\hat{a}^\dagger_{\bs{k}}\hat{a}_{\bs{k}}+\hat{b}^\dagger_{-\bs{k}}\hat{b}_{-\bs{k}})
\end{align*}

Where $\Delta'_{\bs{k}} = \sum_{\delta^{NNN}} \left( e^{i\bs{k}\cdot \bs{\delta}^{NNN}} +  e^{-i\bs{k}\cdot \bs{\delta}^{NNN}}\right)$ and $\tilde{\Gamma}_{\bs{k}} = \frac{1}{2J}\left( \Gamma_{xy}\Delta'_{\bs{k}}-4z\Gamma_z \right)$. Therefore, now:

\begin{equation}
\hat{H}_{\bs{k}} = 2JS\begin{pmatrix} 
z + \Delta_{\bs{k}} + \tilde{\Gamma}_{\bs{k}} & \gamma_{\bs{k}} \\
\gamma_{-\bs{k}} & z - \Delta_{-\bs{k}} + \tilde{\Gamma}_{\bs{k}}
\end{pmatrix}
\end{equation}

We obtain the dispersion relation by changing $z \rightarrow z + \tilde{\Gamma}_{\bs{k}}$ in \ref{DispWithDMI}:

\begin{align}
\hat{H} &= \sum_{\bs{k}} \left( \hat{\alpha}_{\bs{k}}^\dagger \hat{\alpha}_{\bs{k}} \epsilon_{\bs{k}}^+ +\hat{\beta}_{\bs{k}}^\dagger \hat{\beta}_{\bs{k}}  \epsilon_{\bs{k}}^- \right) \label{DispWithDMI} \\
\epsilon_{\bs{k}}^\pm &= 2JS \left( \pm \Delta_{\bs{k}} +  \sqrt{(z + \tilde{\Gamma}_{\bs{k}})^2-\abs{\gamma_{\bs{k}}}^2} \right)
\end{align}